---
title             : "Final Project: The Relationship between Sleep, Depression, Quality of Life, and Socioeconomic Status"
shorttitle        : "Final Project"

author: 
  - name          : "Alexis Adams-Clark"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "1585 E 13th Ave, Straub 339, Eugene, OR 97403"
    email         : "aadamscl@uoregon.edu"
  - name          : "Andrew Fridman"
    affiliation   : "1"
  - name          : "Xi Yang"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "University of Oregon Department of Psychology"

authornote: |

  We would like to acknowledge Daniel Anderson for introducing us to Papaja and our classmates in Introduction to Data Science with R.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "sleep, depression, quality of life, SES"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```


# Methods

## Participants
Participants were recruited through the University of Oregon Human Subjects Pool consisting of undergraduate students enrolled in introductory psychology and linguistics courses. Students received academic credit in exchange for their participants. Participants were not aware of the subject of the study before scheduling their participation, so participants did not self-select into the study. Although they could leave the study after reading informed consent, no participants chose to do so. The University’s Office of Research Compliance (Institutional Review Board) approved this study. 

In total, [#] participants were included in this study, all of who had complete data. [insert demographic characteristics]

## Materials

###Depressive Symptoms Scale

A scale to measure depressive symptoms was created by the researchers for the purposes of this study. Participants were instructed to rate each item on Likert-type scale, where 1 corresponds to “strongly disagree,” and 5 corresponds to “strongly agree.” The scale consisted of three items, including: “Over the last 2 weeks, I have felt little interest or pleasure in doing things;” “Over the last 2 weeks, I have felt down, depressed, or hopeless;” and “Over the last 2 weeks, I have felt tired or had little energy.” Scale items were summed to create a depression score, where higher scores represent higher depressive symptoms.

###Quality of Life Scale
A scale to measure participants’ quality of life was created by the researchers. Participants were instructed to rate each item on Likert-type scale, where 1 corresponds to “strongly disagree,” and 5 corresponds to “strongly agree.” The scale consisted of four items, including: “My life is ideal;” “I am satisfied with my life;” “So far I have been able get the important things I want from life;” and “I have accomplished many of the things in my life.” Scale items were summed to create a total quality of life score, where higher scores represent better quality of life.

###Sleep Quality Scale
A scale to measure participants’ sleep quality was created by the researchers. Participants were instructed to rate each item on Likert-type scale, where 1 corresponds to “strongly disagree,” and 5 corresponds to “strongly agree.” The scale consisted of three items, including: “I am satisfied with the time I spend sleeping;” “I am satisfied with my quality of sleep”; and “When I wake up, I feel refreshed.” Scale items were summed to create a total sleep quality score, where higher scores represent better sleep quality. 


###Demographics Questionnaire
Participants were also asked to complete a demongraphics quationnaire, which asked about their socioeconomic status, race/ethnicity, and educational background. 

## Procedure

[need to insert]

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses. The authors interpreted our correlational results as small (.10-.29), medium (.30-.49), and large correlations (.50-1.00).


# Results

# Discussion


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
